Hardware-in-the-Loop Plant Modeling for Autonomous Vehicle

### 2022-03-26 reportTII-ArticleHIL.tex
Report Update
Updated the NN Model, Modeling Vehicle Subsystems, and Validation & Testing sections.

\usepackage{tikz}
% \usetikzlibrary{calc,tikzmark}
\usetikzlibrary{backgrounds,calc,patterns,decorations.pathmorphing,decorations.markings,shapes,arrows,snakes,tikzmark}

@@

 \section{Modeling Dynamic System using Neural Network}

 In this section, we illustrate how a neural network can be used to model a
 dynamic system in general. Neural network is a system approximation technique
 that .......
 The main structure of a neural network consists neurons, input layer of
 neurons, one or more hidden layers, and the output layer of neurons. The
 neurons of the hidden layers and the output layer have processing capabilities.
 The actual network of a nural network following parallel and distributed system
 architecture. .....

 \missingfigure{Add a generic picture/block diagram of a multiple-input multiple output
   dynamic system}
 \todo[inline]{Add a generic picture of a neural network}
 \todo[inline]{Weights of a neural network}

@@

\section{Modeling Vehicle Subsystems}
To start this project, we first read documentation outlining the uses of
MATLAB's System Identification Toolbox. From there we worked on MATLAB tutorials
on how to model systems from data using System Identification and then find the
most accurate model. We specifically tried to find examples with Multiple-Input
Multiple-Output (MIMO) and Multiple-Input Single-Output (MISO) systems, since
most of the vehicle subsystems we will model fall into one of these categories.
A literature review was also conducted to see how systems with small
non-linearities like our steering subsystem were modeled using System
Identification. On October 7th we traveled to AutonomouStuff in order to collect
data from the steering, acceleration, and braking subsystems in an autonomous
vehicle. The data we collected will be used to generate and then verify our
models.

\subsection{System Requirements}
Based on the system requirements listed above, the plant model will meet the following specifications:
\begin{itemize}
\item The steering subsystem will be modeled first due to its non-linearities, depicted in Fig.~\ref{fig:nonlinGraph}, and because it is important to the operation of the vehicle 
\item The steering subsystem will be able to handle small steering angles
\item Each vehicle subsystem will be modeled separately
\end{itemize}
The vehicle plant model will fulfill the requirements listed below:
\begin{itemize}
    \item The resulting plant model will consist of accurate subsystem models, as defined above
    \item The subsystems can be used to create a HIL testbench
    \item The subsystems can handle very small changes in values accurately
\end{itemize}

All of the subsystems have nonlinear behaviors when there are small changes in the output of the subsystem. The steering subsystem, for example, should behave in a smooth, continuous manner. However, AutonomouStuff observed that when trying to implement features such as lane tracking for the Lexus vehicle platform, that when small changes in the steering angle (less than five degrees) occurred, the torque voltages would momentarily stall and then suddenly change causing a more drastic change. This behavior, depicted in Fig.~\ref{fig:nonlinGraph}, made it very challenging to complete features like lane tracking. These models will aid in the development of controllers that will remove the nonlinearities allowing features like lane tracking to be implemented in a safer and smoother manner.

@@

@@

To start this project, we first read documentation outlining the uses of
MATLAB's System Identification Toolbox. From there we worked on MATLAB tutorials
on how to model systems from data using System Identification and then find the
most accurate model. We specifically tried to find examples with Multiple-Input
Multiple-Output (MIMO) and Multiple-Input Single-Output (MISO) systems, since
most of the vehicle subsystems we will model fall into one of these categories.
A literature review was also conducted to see how systems with small
non-linearities like our steering subsystem were modeled using System
Identification.

@@

On October 7th we traveled to AutonomouStuff in order to collect
data from the steering, acceleration, and braking subsystems in an autonomous
vehicle. The data we collected will be used to generate and then verify our
models. We collected data on the Lexus RX450H vehicle platform shown in
Fig.~\ref{fig:lexusvehicle}. %

\noindent The following is the hardware required for this project:  
	\begin{itemize}
    		\item Laptop
    		\item PACMod ECU
    		\item CANCase 
    		\item CAN bus 
 	\end{itemize}

	\noindent The laptop is used to pass commands or log data, such as steering angle or acceleration or brake pedal position. This data is sent or received using the CANCase and CAN bus. These are connected to the AutonomouStuff designed PACMod ECU, which sends torque voltages to the desired vehicle subsystem allowing the laptop to either control the desired vehicle system or log data.\bigskip

	\noindent The following is the software required for this project: 
	\begin{itemize}
    		\item MATLAB's System Identification Toolbox 
    		\item Vector CANAlyzer 
 	\end{itemize}

\noindent The Vector CANAlyzer software is installed on the laptop and is used to parse the collected data that is sent from the CANCase. This is how we were able to collect logs of data that we would use to develop models of the autonomous vehicle subsystems. MATLAB's System Identification Toolbox is the software that gave us the capabilities to develop these models of the vehicle subsystems. Using this toolbox, we are able to use the logs of data we collected to create sets of estimation and validation data that will be used for training the model.
\\

@@

Using the data we collected for the steering, acceleration, and braking subsystems, we initially separated the data so we could analyze the by-wire and manual modes individually. This effort was made to identify if the models we developed could be used interchangeably regardless of what mode the autonomous vehicle subsystems were operating in. After analyzing the models we developed, we determined that they are not interchangeable as there were differences in the models. When the vehicle subsystem is in by-wire mode, the controller AutonomouStuff developed is generating the torque voltages that are applied to each subsystem motor. As a result, we decided the most accurate model of each subsystem would be developed using the data when the subsystems are in manual mode. For reference, the differences in the models created for by-wire and manual mode are depicted in \ref{sec:simresults} for the steering and acceleration subsystems. For all other models we develop, we will only use manual mode data and will not create models for by-wire mode.
\\

Another decision we made was to start using the Neural Network to create models. We decided to do this instead of continuing with the System Identification Toolbox for multiple reasons. The main one is because while transfer function models are able to accurately model linear systems, they struggle to model highly non-linear systems. Neural Networks work well with highly non-linear systems and are able to accurately model them. Since our systems, and the steering system in particular, exhibit non-linear behavior, they are not able to be accurately modeled by transfer functions. Another reason is that once the System Identification models were created in Simulink, there was confusion over how they would connect to produce the output. Some of the system models, such as the brake system, had four transfer functions due to having four inputs. These models were easily generated, but it wasn’t clear on how they should be connected in order to reflect the physical system. By using Neural Network models this problem can be avoided.

@@

\subsubsection{Brake System}
The brake system was modeled using only the manual log data. It was trained using the data from the first brake log, and the result and error plots can be seen in \autoref{initalBrakeModel}. The plots show that the model is able to predict the outputs relatively accurately, along with keeping the error below 5\%. In order to further verify the accuracy of the model, it was tested using data from the second brake log. The results from that test, shown in \autoref{testedBrakeModel}, also show that the model can accurately predict the output values and do this within the accepted error bound of 5\%.
\begin{figure}[h]
	\centering
	\subcaptionbox{Model Output Tracking and Error Plots \label{initalBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/brake_new_neuralNetworkFig}}
	\subcaptionbox{Model Output Tracking and Error Plots for Log Two Data \label{testedBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/brake_new_neuralNetworkFigLog2Test}}
	\caption{Brake System Training Plots}
\end{figure}

@@

\todo[inline]{Discuss pros and cons of the results you presented in the previous sections on each of the techniques you incorporated}

@@

\todo[inline]{One paragraph on conclusion, make sure you to recommend what modeling technique for the company }
\todo[inline]{One paragraph on future work}

>>>>

\usepackage{tikz}
\usetikzlibrary{shapes,arrows, positioning, quotes}
\usetikzlibrary{arrows.meta, positioning, decorations.markings}
\usetikzlibrary{calc,tikzmark}
\usetikzlibrary{backgrounds,calc,patterns,decorations.pathmorphing,decorations.markings,shapes,arrows,snakes,tikzmark}

@@

 \section{Modeling Dynamic Systems using Neural Networks}

 In this section, we illustrate how a neural network can be used to model a
 dynamic system in general. Neural network is a system approximation technique
 that .......
 The main structure of a neural network consists of neurons, input layer of
 neurons, one or more hidden layers, and the output layer of neurons. The
 neurons of the hidden layers and the output layer have processing capabilities.
 The actual structure of a neural network follows parallel and distributed system
 architecture. .....

 \begin{figure}[htbp]
    \centering
  \begin{tikzpicture}
    \tikzstyle{every node} = [font=\footnotesize]
    \tikzstyle{block} = [draw, rectangle, fill=green!15, rounded corners=5pt, minimum width=1.5 cm, minimum height=1.5 cm]
    % place nodes
    \node[block,text width=1.75cm,text centered](plant){};

    % Connections
	\draw[stealth-,thick]
    (plant.-130)--node[very near end,above](in1){Input 1}++(-1.9,0);    
    \draw[stealth-,thick]
    (plant.180)--node[very near end,above](in2){Input 2}++(-1.9,0);
    %\draw[ultra thick,Dotted] (in2) -- (inn);
    \draw[-stealth,thick]
    (plant.130)--node[very near end,above](inn){Input n}++(1.9,0);

    \draw[-stealth,thick]
    (plant.-45)--node[very near end,above](out1){Output 1}++(1.9,0);
    \draw[-stealth,thick]
    (plant.0)--node[very near end,above](out2){Output 2}++(1.9,0);
    %\draw[ultra thick,Dotted] (out2) -- (outm);
    \draw[-stealth,thick]
    (plant.0)--node[very near end,above](outm){Output m}++(1.9,0);

  \end{tikzpicture}
    \caption{Multiple Input Multiple Output Dynamic System Block Diagram}
    \label{fig:mimoDynamicBlockDiagram}
 \end{figure}

 Figure~\ref{fig:nnActor} shows a generic actor neural network structure. ...
 \begin{figure}
  \centering
      \tikzset{%
        input neuron/.style={
          circle,
          fill=green!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \tikzset{%
        hidden neuron/.style={
          circle,
          fill=blue!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \tikzset{%
        output neuron/.style={
          circle,
          fill=red!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \begin{tikzpicture}[x=1.5cm, y=1.5cm]

        \foreach \m/\l [count=\y] in {1,2,missing,3}
        \node [input neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

        \foreach \m [count=\y] in {1,2,3,missing,4}
        \node [hidden neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2.5-\y*1.1) {};

        \foreach \m [count=\y] in {1,missing,2}
        \node [output neuron/.try, neuron \m/.try] (output-\m) at (4,2.5-\y*1.2) {};

        \foreach \l [count=\i] in {1,2,n}
        \draw [<-] (input-\i) -- ++(-1,0)
        node [above, midway] {$\mathbf{e}_{k,\l}$};

        \foreach \l [count=\i] in {1,2,3,\mbox{$\bar{n}$}}
        \node [above] at (hidden-\i.north) {$\sigma_{\l}$};

        \foreach \l [count=\i] in {1,m}
        \draw [->] (output-\i) -- ++(1,0)
        node [above, midway] {$\hat{u}_{a,k}^{(\l)}$};

        \foreach \i in {1,...,3}
        \foreach \j in {1,...,4}
        \draw [->] (input-\i) -- (hidden-\j);

        \foreach \i in {1,...,4}
        \foreach \j in {1,2}
        \draw [->] (hidden-\i) -- (output-\j);

        \foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
        \node [align=center, above] at (\x*2,1.9) {\l \\ layer};

        \node [align=center, below] at (3,-2.5) {Weight \\ $\mathbf{W}_a$};
        \draw[->,ultra thick](3,-2.5) -- (3,1.6);

      \end{tikzpicture}
  \caption{Actor neural network structure for approximating control input.}
  \label{fig:nnActor}
\end{figure}

 \todo[inline]{Weights of a neural network}

@@

\section{Modeling Vehicle Subsystems}

To start the project, we researched methods of plant modeling for autonomous vehicles and more generically, nonlinear systems. From our research, we found a few methods of modeling that we employed to discover the most accurate representation of each vehicle subsystem. The first method that we tried was developing different system types using Mathworks' MATLAB System Identification Toolbox. Using data we collected from the Lexus vehcile platform at AutonomouStuff,this toolbox allowed us to add input and output data for each vehicle subsystem. With this data, we used one subset to train the system and another subset for validation. Using this toolbox, we tried developing state-space models, NARX models, and transfer function models. We also used Mathworks' MATLAB Neural Network Time Series toolbox to develop a model of each vehicle subsystem using neural network modeling making use of the Bayesian Regularization training method. The following requirements were used to evaluate the effectiveness of our models to determine which plant model should be presented to AutonomouStuff.

\subsection{System Requirements}

In order to accurately model each vehicle subsystem so that safe, reliable controllers can be developed for autonomous control, the following criteria will be met:
\begin{itemize}
\item Each vehicle subsystem will be able to track nonlinearities depicted in Fig.~\ref{fig:nonlinGraph} associated with small changes in the output of each subsystem
\item Each vehicle subsystem plant model will track any desired inputs within the specified error bounds
	\begin{itemize}
		\item Steering Angle within 5 degrees
		\item Pedal Positions within 5\%
	\end{itemize}
\item Each vehicle subsystem will be modeled independently from other vehicle subsystems
\end{itemize}

The vehicle plant model will fulfill the requirements listed below:
\begin{itemize}
    \item The resulting plant model will consist of accurate subsystem models, as defined above
    \item The subsystems can be used to create a HIL testbench
    \item The subsystems can handle very small changes in their output accurately
\end{itemize}

All of the subsystems have nonlinear behaviors when there are small changes in the output of the subsystem. The steering subsystem, for example, should behave in a smooth, continuous manner. However, AutonomouStuff observed that when trying to implement features such as lane tracking for the Lexus vehicle platform, that when small changes in the steering angle (less than five degrees) occurred, the torque voltages would momentarily stall and then suddenly change causing a more drastic change. This behavior, depicted in Fig.~\ref{fig:nonlinGraph}, made it very challenging to complete features like lane tracking. These models will aid in the development of controllers that will remove the nonlinearities allowing features like lane tracking to be implemented in a safer, smoother manner.

@@

\subsection{Steering System Modeling}

\subsection{Acceleration System Modeling}

\subsection{Brake System Modeling}

@@

%To start this project, we first read documentation outlining the uses of
%MATLAB's System Identification Toolbox. From there we worked on MATLAB tutorials
%on how to model systems from data using System Identification and then find the
%most accurate model. We specifically tried to find examples with Multiple-Input
%Multiple-Output (MIMO) and Multiple-Input Single-Output (MISO) systems, since
%most of the vehicle subsystems we will model fall into one of these categories.
%A literature review was also conducted to see how systems with small
%non-linearities like our steering subsystem were modeled using System
%Identification.

@@

We traveled to AutonomouStuff in order to collect
data from the steering, acceleration, and braking subsystems in an autonomous
vehicle. The data we collected will be used to generate and then verify our
models. We collected data on the Lexus RX450H vehicle platform shown in
Fig.~\ref{fig:lexusvehicle}. %

\noindent The following is the hardware components required to collect the data:  
	\begin{itemize}
    		\item Laptop
    		\item PACMod ECU
    		\item CANCase 
    		\item CAN bus 
 	\end{itemize}

\noindent The laptop is used to pass commands or log data, such as steering angle or acceleration or brake pedal position. This data is sent or received using the CANCase and CAN bus. These are connected to the AutonomouStuff designed PACMod ECU, which sends torque voltages to the desired vehicle subsystem allowing the laptop to either control the desired vehicle system or log data. The Vector CANAlyzer software is installed on the laptop and is used to parse the collected data that is sent from the CANCase. This is how we were able to collect logs of data that we would use to develop models of the autonomous vehicle subsystems.
\\

@@

Using the data we collected for the steering, acceleration, and braking subsystems, we initially separated the data so we could analyze the by-wire and manual modes individually. This effort was made to identify if the models we developed could be used interchangeably regardless of what mode the autonomous vehicle subsystems were operating in. After analyzing the models we developed, we determined that they are not interchangeable as there were differences in the models. When the vehicle subsystem is in by-wire mode, the controller AutonomouStuff developed is generating the torque voltages that are applied to each subsystem motor. As a result, we decided the most accurate model of each subsystem would be developed using the data when the subsystems are in manual mode. For reference, the differences in the models created for by-wire and manual mode are depicted in \ref{sec:simresults} for the steering and acceleration subsystems. For all other models we develop, we will only use manual mode data and will not create models for by-wire mode.

@@

\subsubsection{Steering System}

The following are the results of neural network training for the steering system.

\begin{figure}[htbp]
	\centering
	\subcaptionbox{Model Output Tracking and Error Plots for Data Subset 1 \label{initalBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/steeringNeuralNetworkTrainedOutput}}
	\subcaptionbox{Model Output Tracking and Error Plots for Data Subset 2 \label{testedBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/steeringNeuralNetworkTrainedOutput}}
	\caption{Steering System Training Plots}
\end{figure}

\subsubsection{Acceleration System}

The following are the results of neural network training for the acceleration system.

\begin{figure}[htbp]
	\centering
	\subcaptionbox{Model Output Tracking and Error Plots for Data Subset 1 \label{initalBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/accelNeuralNetworkTrainedOutput}}
	\subcaptionbox{Model Output Tracking and Error Plots for Data Subset 2 \label{testedBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/accelNeuralNetworkTrainedOutput2}}
	\caption{Acceleration System Training Plots}
\end{figure}

\subsubsection{Brake System}

The brake system was modeled using only the manual log data. It was trained using the data from the first brake log, and the result and error plots can be seen in \autoref{initalBrakeModel}. The plots show that the model is able to predict the outputs relatively accurately, along with keeping the error below 5\%. In order to further verify the accuracy of the model, it was tested using data from the second brake log. The results from that test, shown in \autoref{testedBrakeModel}, also show that the model can accurately predict the output values and do this within the accepted error bound of 5\%.
\begin{figure}[h]
	\centering
	\subcaptionbox{Model Output Tracking and Error Plots \label{initalBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/brake_new_neuralNetworkFig}}
	\subcaptionbox{Model Output Tracking and Error Plots for Log Two Data \label{testedBrakeModel}}
		{\includegraphics[width=0.48\linewidth]{figs/img/brake_new_neuralNetworkFigLog2Test}}
	\caption{Brake System Training Plots}
\end{figure}

@@

For each vehicle subsystem, the results show that the neural network models were the most accurate models. Other models that we developed, such as the transfer function models, were relatively accurate when tracking the data, but those models were consistently falling outside of the error bounds that were defined at the start of the project. The neural network models, however, track the data more accurately and reduced the error significantly so that all error requirements were met. One downside to using neural networks as the plant model is it causes some complexities when trying to include these models into the HIL bench software for use in developing and testing controllers.

@@
